<!--
%\VignetteEngine{knitr}
%\VignetteIndexEntry{List filtering}
-->

```{r, echo = FALSE, message = FALSE}
library(rlist)
knitr::opts_chunk$set(comment="#",error=FALSE,tidy=FALSE)
```

# List filtering

List filtering is to select list members that meet given criteria. In rlist package, more than ten functions are related with list filtering. Basically, they all perform a mapping but aggregate the results in different ways.

```{r}
devs <- 
  list(
    p1=list(name="Ken",age=24,
      interest=c("reading","music","movies"),
      lang=list(r=2,csharp=4,python=3)),
    p2=list(name="James",age=25,
      interest=c("sports","music"),
      lang=list(r=3,java=2,cpp=5)),
    p3=list(name="Penny",age=24,
      interest=c("movies","reading"),
      lang=list(r=1,cpp=4,python=2)))
```

## list.filter

`list.filter` filters a list by an expression that returns a logical value. Only the list members for which the value of that expression turns out to be `TRUE` will be part of the results.

```{r}
str(list.filter(devs, age >= 25))
str(list.filter(devs, "sports" %in% interest))
str(list.filter(devs, mean(as.numeric(lang)) >= 3))
```

Meta-symbols defined in lambda expression can also be used. The following code will pick up the list member whose index is even.

```{r}
str(list.filter(devs, .i %% 2==0))
```

## list.find

In some cases, we don't need to find all the results given the criteria. Rather, we only need to find a few, sometimes only one. `list.find` will avoid searching across all list member but stops at a specific number of items found.

```{r}
list.find(devs, age >= 25, 1)
```

## list.findi

`list.findi` only returns the index of the members found.

```{r}
list.findi(devs, age >= 23, 2)
```

## list.takeWhile

`list.takeWhile` keeps taking members while a condition holds true.

```{r}
list.takeWhile(devs, lang$r >= 2)
```

## list.skipWhile

`list.skipWhile` keeps skipping members while a condition holds true.

```{r}
list.skipWhile(devs, lang$r <= 2)
```

## list.if

`list.if` returns a logical vector that indicates whether a condition holds for each member of a list.

```{r}
list.if(devs, "music" %in% interest)
list.if(devs, "java" %in% names(lang))
```

## list.which

`list.if` returns a integer vector of the indices of the members of a list that meet a given condition.

```{r}
list.which(devs, "music" %in% interest)
list.which(devs, "java" %in% names(lang))
```

## list.all

`list.all` returns `TRUE` if all the members of a list satisfy a given condition, or `FALSE` otherwise.

```{r}
list.all(devs, mean(as.numeric(lang)) >= 3)
list.all(devs, "r" %in% names(lang))
```

## list.any

`list.any` returns `TRUE` if at least one of the members of a list satisfies a given condition, or `FALSE` otherwise.

```{r}
list.any(devs, mean(as.numeric(lang)) >= 3)
list.any(devs, "python" %in% names(lang))
```

## list.count

`list.count` return a scalar integer that indicates the number of members of a list that satisfy a given condition.

```{r}
list.count(devs, mean(as.numeric(lang)) >= 3)
list.count(devs, "r" %in% names(lang))
```

## list.match

`list.match` filters a list by matching the names of the list members by a regular expression pattern.

```{r}
list.match(devs, "p[12]")
```

## list.remove

`list.remove` removes list members by index or name.

```{r}
list.remove(devs, c("p1","p2"))
list.remove(devs, c(2,3))
```

## list.exclude

`list.exclude` removes list members that satisfy given condition.

```{r}
list.exclude(devs, "sports" %in% interest)
```

## subset

`subset` is implemented for list object in a way that combines `list.filter` and `list.map`. This function basically filters a list while at the same time maps the qualified list member by an expression.

```{r}
subset(devs,age>=24,name)
subset(devs,"reading" %in% interest, sum(as.numeric(lang)))
```

