<!--
%\VignetteEngine{knitr}
%\VignetteIndexEntry{List grouping}
-->

```{r, echo = FALSE, message = FALSE}
library(rlist)
knitr::opts_chunk$set(comment="#",error=FALSE,tidy=FALSE,size=1.5)
```

# List grouping

rlist supports multiple types of grouping. 

```{r}
devs <- 
  list(
    p1=list(name="Ken",age=24,
      interest=c("reading","music","movies"),
      lang=list(r=2,csharp=4,python=3)),
    p2=list(name="James",age=25,
      interest=c("sports","music"),
      lang=list(r=3,java=2,cpp=5)),
    p3=list(name="Penny",age=24,
      interest=c("movies","reading"),
      lang=list(r=1,cpp=4,python=2)))
```

## list.group

`list.group` is used to classify the list members into different groups by evaluating a given expression and see what value it takes. The expression often produces a scalar value such as a logical value, a character value, and a number. Each group denotes a unique value that expression takes for at least one list member, and all members are put into one and only one group.

```{r}
str(list.group(devs, age))
str(list.group(devs, length(interest)))
```

## list.ungroup

`list.group` will produce a nested list in which the first level are groups and the second level are the original list members put into different groups. 

`list.ungroup` is used to revert this process, or eradicate the group level of a list.

```{r}
ageGroups <- list.group(devs, age)
str(list.ungroup(ageGroups))
```

## list.cases

In non-relational data structures, a field can be a vector of multiple values. `list.cases` is used to find out all possible cases by evaluating a vector-value expression for each list member.

In data `devs`, field `interest` is usually a character vector of mutiple values. The following code will find out all possible interests for all list members.

```{r}
list.cases(devs, interest)
```

Or use similar code to find out all programming languages the developers use.

```{r}
list.cases(devs, names(lang))
```

## list.class

To group by cases, use `list.class` function. This function basically classify all list members case by case. Therefore, a long and nested list will be produces, in which the first-level denotes all the cases, and the second-level includes the original list members.

Note that each list member may belong to mulitple cases, therefore the classification of the cases for each member is not exclusive. You may find one list member belong to multiple cases in the resulted list.

Case classification by `interest`:

```{r}
str(list.class(devs, interest))
```

Case classification by `names(lang)`:

```{r}
str(list.class(devs, names(lang)))
```

