<!--
%\VignetteEngine{knitr}
%\VignetteIndexEntry{List joining}
-->

```{r, echo = FALSE, message = FALSE}
library(rlist)
knitr::opts_chunk$set(comment="#",error=FALSE,tidy=FALSE,size=1.5)
```

# List joining

rlist provides `list.join` to perform simple joining of two lists, and `list.merge` to perform modification merge of a series of lists.

```{r}
devs <- 
  list(
    p1=list(name="Ken",age=24,
      interest=c("reading","music","movies"),
      lang=list(r=2,csharp=4,python=3)),
    p2=list(name="James",age=25,
      interest=c("sports","music"),
      lang=list(r=3,java=2,cpp=5)),
    p3=list(name="Penny",age=24,
      interest=c("movies","reading"),
      lang=list(r=1,cpp=4,python=2)))
```


## list.join

`list.join` is used to join two lists by a key evaluated from either a common expression for the two lists or two separate expressions for each list.

```{r}
newinfo <-
  list(
    p1=list(name="Ken",email="ken@xyz.com"),
    p2=list(name="Penny",email="penny@xyz.com"),
    p3=list(name="James",email="james@xyz.com"))
str(list.join(devs,newinfo,name))
```

## list.merge

`list.merge` is used to recursively merge a series of lists with the later always updates the former. It works with two lists, as shown in the example below, in which a revision is merged with the original list.

```{r}
rev1 <-
  list(
    p1=list(age=25),
    p2=list(lang=list(r=2,cpp=4)),
    p3=list(lang=list(r=2,python=NULL)))
str(list.merge(devs,rev1))
```

The function also works with multiple lists. When the second revision is obtained, the three lists can be merged in order.

```{r}
rev2 <-
  list(
    p1=list(lang=list(csharp=5)),
    p2=list(age=24,lang=list(r=3)))
str(list.merge(devs,rev1,rev2))
```

