<!--
%\VignetteEngine{knitr}
%\VignetteIndexEntry{List}
-->

```{r, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(comment="#",error=FALSE,tidy=FALSE)
```

# List environment

List environment is an alternative construct designed for easier command chaining. `List()` function wraps a list within an environment where almost all functions in this package are defined but result in the next List environment for further operations.

Suppose we work with the following list.

```{r}
devs <- 
  list(
    p1=list(name="Ken",age=24,
      interest=c("reading","music","movies"),
      lang=list(r=2,csharp=4,python=3)),
    p2=list(name="James",age=25,
      interest=c("sports","music"),
      lang=list(r=3,java=2,cpp=5)),
    p3=list(name="Penny",age=24,
      interest=c("movies","reading"),
      lang=list(r=1,cpp=4,python=2)))
```

To create a `List environment`, run

```{r}
m <- List(devs)
```

then we can operate the environment-based object `m` with `map()`, `filter()` and other functions, or extract the inner data with `m$data`. All inner functions return `List environment`, which facilities command chaining.

For example, map each member to their name.

```{r}
m$map(name)
```

Note that the resulted object is also a `List environment` although its printed results include the inner data. To use the result with external functions, we need to extract the inner data by calling `m$data`.

Get all the possible cases of interests for those whose R experience is longer than 1 year.

```{r}
m$filter(lang$r > 1)$
  cases(interest)$
  data
```

Calculate an integer vector of the average number of years using R for each interest class.

```{r}
m$class(interest)$
  map(case -> length(case))$
  call(unlist)$
  data
```

