<!--
%\VignetteEngine{knitr}
%\VignetteIndexEntry{List searching}
-->

```{r, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(comment="#",error=FALSE,tidy=FALSE)
```

# List searching

rlist provides searching capabilities, that is, find values within a list without specifying the key or searching path. `list.search` handles a variety of search demands.

## Exact search

Exact search is to find values only by logical examinations. Suppose we search the following list.

```{r}
x <- list(p1 = list(type="A",score=c(c1=9)),
  p2 = list(type=c("A","B"),score=c(c1=8,c2=9)),
  p3 = list(type=c("B","C"),score=c(c1=9,c2=7)),
  p4 = list(type=c("B","C"),score=c(c1=8,c2=NA)))
```

### Search identical values

First, we search all values in the list that is identical to "A".

```{r}
list.search(x, identical, "A")
```

Only values that are identical to character vector `"A"` will be put in the resulting list. We can also unlist the result.

```{r}
list.search(x, identical, "A", unlist = TRUE)
```

Then, we search all values identical to `c("A","B")`.

```{r}
list.search(x, identical, c("A","B"))
```

Next, we search if there is a numeric vector `c(10,8)`.

```{r}
list.search(x, identical, c(10,8))
```

The result is none. If you are familiar with how function `identical` works, you should not feel surprises since this may be the strongest comparer to tell whether two objects are the same: Two objects are identical when and only when they have absolutely the same structure including values and names. That explains why there is no numeric vector like `c(10,8)` because all numeric vectors in `x` are named vectors like `c(c1=10,c2=8)`. 

To compare values between atomic vectors just like using `==`, we can use `allEqual`, `anyEqual`, `allUnequal`, or `anyUnequal` as the comparer function.

### Seach all equal values

Search all values equal to 9.

```{r}
list.search(x, allEqual, 9)
```

Search all values equal to `c(8,9)`.

```{r}
list.search(x, allEqual, c(8,9))
```

Search all values equal to `c(8,9)` ignoring `NA`.

```{r}
list.search(x, allEqual, c(8,9), na.rm = TRUE)
```

### Search any equal values

Search all character vectors in which any element equals "A".

```{r}
list.search(x, anyEqual, "A")
```


Search all numeric vectors in which any element equals 8.

```{r}
list.search(x, anyEqual, 8)
```


## Fuzzy search

The comparison is flexible enough to support fuzzy searching using functions provided by [`stringdist`](http://cran.r-project.org/web/packages/stringdist/index.html) package. Consider the following list.

```{r}
x <- list(
    p1 = list(name="Ken",age=24),
    p2 = list(name="Kent",age=26),
    p3 = list(name="Sam",age=24),
    p4 = list(name="Keynes",age=30),
    p5 = list(name="Kwen",age=31))
```

rlist's built-in functions `anyLike(dist,...)` and `allLike(dist,...)` that internally call `stringdist::ain` and `stringdist::stringdist` respectively handle fuzzy search to meet a wide range of demands. 

For both functions, `dist` means the maximum string distance between the actual value in the list and the search term you specify, and `...` is the additional parameters passed to their internally called functions. More specifically,

- `allLike(1)` means the string distances of **ALL** values in the character vector must be no greater than one to be qualified as a search result.
- `anyLike(1)` means that **ANY** value in the character vector in the list whose string distance is no greater than one will make the vector qualified as a search result.

Similar logic also applies to `allUnlike()` and `anyUnlike()` as their names suggests.

For example, if we want to find out names similar with `"Ken"` with maximum distance 1, we only have to specify `fun = allLike(1)`.

```{r}
list.search(x, allLike(1), "ken", unlist = TRUE)
```

`anyLike(dist)` is the counterpart of `allLike(dist)` functioning similar with `%in%` operator. It tells whether a character vector includes a similar value within a specific range of distance.

Suppose we are working with the following data.

```{r}
x <- list(
    p1 = list(name=c("Ken", "Ren"),age=24),
    p2 = list(name=c("Kent", "Potter"),age=26),
    p3 = list(name=c("Sam", "Lee"),age=24),
    p4 = list(name=c("Keynes", "Bond"),age=30),
    p5 = list(name=c("Kwen", "Hu"),age=31))
```

Search all character vectors in which any element is like "Ken" within string distance 1.

```{r}
list.search(x, anyLike(1), "Ken")
```

Search all character vectors in which all elements are unlike "Ken" due to string distance no less than 2.

```{r}
list.search(x, allUnlike(2), "Ken")
```

## Fuzzy filtering

The fuzzy search functions also work with filtering functions.

Consider the following data.

```{r}
x <- list(
    p1 = list(name=c("Ken", "Ren"),age=24),
    p2 = list(name=c("Kent", "Potter"),age=26),
    p3 = list(name=c("Sam", "Lee"),age=24),
    p4 = list(name=c("Keynes", "Bond"),age=30),
    p5 = list(name=c("Kwen", "Hu"),age=31))
```

We can also use fuzzy search comparers with `list.filter`. For example, filter all list members whose `name` has any character value like `Ken` with maximum distance 1, and output their pasted names as a named character vector. Here we use pipeline.

```{r}
library(pipeR)
x %>>%
  list.filter(name %>>% anyLike(1)("Ken")) %>>%
  list.mapv(paste(name,collapse = " "))
```
