<!--
%\VignetteEngine{knitr}
%\VignetteIndexEntry{List sorting}
-->

```{r, echo = FALSE, message = FALSE}
library(rlist)
knitr::opts_chunk$set(comment="#",error=FALSE,tidy=FALSE)
```

# List sorting

rlist package provides functions for sorting list members by lambda expression.

```{r}
devs <-
  list(
    p1=list(name="Ken",age=24,
      interest=c("reading","music","movies"),
      lang=list(r=2,csharp=4,python=3)),
    p2=list(name="James",age=25,
      interest=c("sports","music"),
      lang=list(r=3,java=2,cpp=5)),
    p3=list(name="Penny",age=24,
      interest=c("movies","reading"),
      lang=list(r=1,cpp=4,python=2)))
```

## list.order

`list.order` evaluates the given lambda expressions and find out the order by default ascendingly. If the values for some members tie, the next values of the next expression, if any, will count.

To adjust the order, use `desc` function or simple write a minus operator (`-`) before the expression.

```{r}
list.order(devs, age)
list.order(devs, length(interest))
list.order(devs, desc(lang$r))
list.order(devs, max(unlist(lang)))
list.order(devs, desc(length(interest)), desc(lang$r))
```

## list.sort

`list.sort` produces a sorted list of the original list members. Its usage is exactly the same as `list.order`.

```{r}
str(list.sort(devs, age))
str(list.sort(devs, length(interest)))
str(list.sort(devs, desc(lang$r)))
str(list.sort(devs, max(unlist(lang))))
str(list.sort(devs, desc(length(interest)), desc(lang$r)))
```
